Algebraic Groups\vspace{-1ex}

### 2019-11-04 Algebraic Geometry/Algebraic-Geometry.tex
update

>>>>

\section{November 4th, 2019}
We're going to talk a bit more about projective varieties because we really haven't done a lot of examples and that is a shame.

\subsection{What we want to do}
By the end of the quarter (we have four more weeks, but only three more weeks worth of meetings with holidays), we want to 
have talked a bit about sheaves and dimension/intersection theory. Continuing into next quarter, we are going to dip into schemes, although perhaps not immediately.

\subsection{Some tangible examples}
One idea is to read Harris' \textit{Algebraic Geometry: a first course}. Now we're talking to talk about $\P n$.

First things first, we form $\P n$ as a set as the lines in $k^{n+1}$ or $k^{n+1}\setminus \{0\}/k^\times$. A point in $\P n$ 
has homogeneous coordinates $[x_0:\cdots:x_n]$, which is only well-defined up to scaling. But if any $x_i$ is nonzero, we can normalize this 
coordinate to get well-defined coordinates. The collection of sets $\{x_i\ne 0\}=\A n\subset \P n$ is called the standard affine open 
cover of $\P n$.

A thought: if we have an affine variety $Z\subseteq \A n\subseteq \P n$, how can we view the closure of $Z$ in $\P n$?
Given $f\in k[x_0,\dots,x_n]$ homogeneous, $Z(f)\subseteq \P n$ is well-defined. More generally, if $I$ is a homogeneous ideal (generated by homogeneous polys)
$Z(I)$ makes sense and furthermore $Z((f_1,\dots,f_k))=\cap_i Z(f_i)$. It ends up that these are precisely the closed sets in the Zariski topology.

\begin{thm}[Projective Nullstellensatz]
	$I(Z(J))=\sqrt{J}$ for all homogeneous $J$ excipt $J$ such thtat $\sqrt{J}$ is the irrelevant ideal $(x_0,\dots,x_n)$.
\end{thm}

\subsection{Homogenization/dehomogenization}
Given a homogenous polynomial $f(x_0,\dots,x_n)$ of degree $d$, we can \textit{dehomogenize} things by setting $x_0=1$. For instance $x_0^2+x_1^2+x_2^2\mapsto 1+x_1^2+x_2^2$.
Given a polynomial $g(x_1,\dots,x_n)$, we just look at 
\[\tilde g(x_0,\dots,x_n)=x^{\deg g}g(x_1/x_0,\dots,x_n/x_0).\]
One can see relatively easily that $\tilde g$ is a homogeneous polynomial of degree $d$.

\begin{prob}
	What happens when you homogenize then dehomogenize? How about the other way around?
\end{prob}

### 2019-11-04 Algebraic Groups/Algebraic-Groups.tex
update

>>>>

\section{November 4th, 2019}
Today we are going to do a bit of review as well as do a little more work with $G$-torsors.

\subsection{Review}
Recall the following setup: $S=\Spec A$ is Noetherian. Then let $G\to S$ be an affine group scheme of finite type over $S$. Let $P\to S$ 
be a scheme with a right $G$-action:
\[P\times_S G\xrightarrow{\sigma}P.\]
\begin{defn}\label{defn:torsor}
	Then $P\to S$ is a \textbf{$G$-torsor} if either of the following hold:
	\begin{itemize}
		\item $P\to S$ is affine, surjective, of finite type, and furthermore transitive (that is $P\times_S G\xrightarrow{(\sigma,\pi_1)}P\times_SP$ is an isomoporphism).
		\item There exists a faithfully flat $S'\to S$ with $P\times_S S'\cong G\times_S S'$ withere this isomorphism is $G\times_S S'$-equivariant.
	\end{itemize}
\end{defn}
\begin{rmk}
	If $G$ is a linear algebraic gorup over $k$ and $S$ is defined over $k$, then we say $P\to S$ is a $G$-torsor if it is a torsor under $G\times_k S$.
\end{rmk}

Now a \textbf{trivial $G$-torsor} over $S$ is $S\times G\to S$ if and only if $P\to S$ has a section.

Some examples:
\begin{itemize}
	\item The $\Gm$ torsors over $S$ are in bijection with the line bundles over $S$.
	\item The $\GL_n$ torsors over $S$ are in bijection with the vector bundles over $S$.
\end{itemize}
Since line and vector bundles are tivialized in the Zariski topology, we see that the second condition in definition~\ref{defn:torsor} can be replaced with: 
There exists an open cover $\{S_i\}$ such that $P|_{S_i}\cong S_i\times G$.

We proved some stuff with descent. Look at it.
\begin{cor}
There exists an equivalence between $\Alg_A$ and the category of $B$ algebras $C$ along with isomorphisms $\alpha:p_1^\ast C\to p_2^\ast C$ along with the cocyle condition.
\end{cor}
\begin{rmk}
	To prove this, we descend a $(B,\alpha)$ to an $A$-module and show that the multiplication on $B$ descends nicely to $A$.
\end{rmk}
\begin{cor}
	We can also descend $G$-torsors. That is, we have an equivalence between $G$-torsors $P\to S$ and $G$-torsors $P'\to S'$ with isomorphisms and the cocycle condition.
\end{cor}
This last result takes quite a bit of doing although we have all the machinery we need. This is the 
one we're going to be using.

\subsection{New Stuff}
Let $X$ be a scheme. Then $\operatorname{Pic}(X)$ is the group of line bundles, or equivalently $\Gm$-torsors on $X$.
\begin{thm}
	$\operatorname{Pic}(X)=H^1(X,\O_X^\times)=H^1(X,\Gm)$.
\end{thm}
Notice that $\O_X^\times$ is a sheaf assigning to each $U$ $\O_X(U)^\times$. On Zariski opens, this is the same as $\Gm(U)$.

To prove the above result, we need to discuss 
\subsection{\texorpdfstring{\v C}{C}ech Cohomology}
Assume $X$ is separated. Let $\calU=\{U_i\}$ be an affine covering where $U_i\cap U_j$ is also affine. Then considering the complex 
\[\sqcup U_i\cap U_j\cap U_k\to \sqcup U_i\cap U_j\to \sqcup U_i\to X\]
where we have several a parallel maps for each $n-1$-tuple of intersections in the previous term. Then we can applie $\O_X^\times$:
\[\prod \O_X(U_i)^\times\to \prod\O_X(U_i\cap U_j)^\times\to\cdots\]
where, for instance, the first map above sends a product of maps $(s_i)$ to $(s_i|_{U_i\cap U_j}-s_j|_{U_i\cap U_j})$.

\begin{defn}
	$\hat H_\calU^i(X,\O_X^\times)$ is the $i^{th}$ cohomoogy of the above complex.
\end{defn}
Then it can be shown that $\hat H_\calU^i(X,\O_X^\times)$ is independent of cover if the $U_i$ are affine.

So what is $H^1(X,\O_X^\times)$? It is the $s_{i,j}in\O_X^\times(U_i\cap U_j)$ modulo the cocycle condition.

\subsection{A new result}
\begin{thm}
	$H^1(X,G)$ is identified with $G$-torsors.
\end{thm}
What is $H^1(X,G)$? We are going to try to recover \v Cech cohomology. Let $X$ be quasi-compact and let $\cup_1^n U_i\to X$ be 
faithfully flat. The seperatedness gets us the intersections are affine.

Then we can play the same game, but this time applying $G$: (notice that $U_i\cap U_j=U_i\times_X U_j$)
\[\prod G(U_i)\to \prod G(U_i\times_X U_j)\to\cdots\]
where (notice now $G$ may be nonabelian!) we map 
\[(s_i)\mapsto (s_i|_{U_i\times_X U_j}\cdot s_j|_{U_i\times_X U_j}^{-1})_{i,j}\]
and 
\[(s_{i,j})\mapsto (s_{i,j}\cdot s_{j,k}\cdot s_{i,k}^{-1})\]
where each $s$ is restricted to $U_i\times_XU_j\times_X U_k$.

Then we can define $\hat H_\calU^1(X,G)$ to be the first cohomology of the above chain. Then 
\begin{defn}
	The \textbf{flat cohomology} is 
	\[H^1_\text{flat}(X,G)=\operatorname{colim}_\calU H^1_\calU(X,G).\]
\end{defn}

The overall result here is 
\begin{thm}
	The $G$-torsors on $X$ in bijection with the elements of $H^1_\text{flat}(X,G)$.
\end{thm}
